\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-13 forwards-and-backwards.tex
added MMT

The lookdown processes have been carefully constructed so that
observations about the past spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels.
In other words, the dynamics of the lookdown
process preserve the conditionally uniform (or in the limit, conditionally Poisson) structure
-- if started with uniform levels,
levels are uniform at all future times.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2019}
tells us that by ``removing labels'' from the lookdown process $\lp$
we recover the population process $\eta$.
% In particular, for any
% solution to the martingale problem of Definition~\ref{defn:mgale_construction}
% % of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
% there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
% solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
% for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
% and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
% are independently and uniformly distributed on $[0, N]$.
% Similarly, in the scaling limit:
% for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
% Theorem~\ref{nonlocal_convergence}
% %there is a lookdown process $(\lp^N_t)_{t \ge 0}$
% there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
% for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
% and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
% $\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
% Furthermore, convergence of the lookdown processes would imply convergence of the population processes
% \citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

@@

\begin{proof}[of Proposition~\ref{thm:mmt_application}]
    This proposition is the content of the Markov Mapping Theorem,
    Theorem A.2 in \citet{etheridge/kurtz:2019},
    applied to our situation.
    \comment{TODO: copy the MMT into our appendix and refer to it}
    In the notation of that theorem, $\gamma$ is what we have called $\kappa$ above,
    $\alpha$ is the transition function that assigns levels uniformly on $[0, N]$ (in the first case)
    or as a Poisson process with Lesbegue intensity (in the limiting case),
    and we need a continuous $\psi^N(\lp)$ such that $|A^N f(\lp)| \le c_f \psi^N(\lp)$
    for all $f$ in the domain of $A^N$ (and similarly a function $\psi$ for $A$).
    We also need that applying the lookdown generator to a function and averaging over levels
    is equivalent to applying the population process generator to the function
    whose dependence on levels has been averaged out,
    a condition which we precisely state, and verify,
    in Appendix~\ref{sec: Markov Mapping Theorem Application}.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C (1 + u |F(x, \eta)|), \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the most troublesome term in $Af(\lp)$,
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align*} % \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align*}
    The first line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the second line provides us with the required bound:
    we absorb $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and take $\psi(\lp) = \langle F(x, \eta) u e^{-u}, \lp(dx, du) \rangle$.
    % Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    % for a given $\lp$ (and hence $\eta = \kappa(\lp)$),
    % $\sup_x |F(x, \eta)|$ is finite.
    % Therefore, we may bound \eqref{eqn:part_of_psi} by
    % $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    % To do this we must restrict $\lp$ to lie in the space of measures
    % for which this is finite;
    % however, this is a completely metrizable space
    % \citep[see Condition 2.1 in][]{etheridge/kurtz:2018},
    % and our processes fall in this space (by conditional uniformity).
\end{proof}

@@

\comment{TODO copy in the MMT and refer to it earlier}

% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %%
\section{Markov Mapping Theorem}

\comment{TODO copy in the MMT and refer to it earlier}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{OLDER STUFF Results on Lookdown Representations}
    \label{sec: Markov Mapping Theorem Application}

% % % % % % % % % % % % %
\subsubsection{Convergence of generators for the lookdown process}
    \label{sec:lookdown_generator_proofs}

\comment{
    I did this informal writeup of convergence of the lookdown generator
    without looking at the previous version
    mostly to double-check what we have there
    (which is probably better).
}

In Definition~\ref{defn:lookdown_mgale} we saw that the generator
of $(\lp_t)_{t \ge 0}$ is the sum of two parts,
equations \eqref{eqn:birth_generator} and \eqref{eqn:level_generator}.
Assume that there is a $u_*$ such that $g(x, u) = 1$ for all $u \ge u_*$.
We consider these two in turn.

First, the contribution from births is
\begin{align*}
\begin{split}
f(\lp)
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 N}
        \int_u^N
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{N}
        \int_u^N \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align*}
We have that
\begin{align*}
    &\theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy) \\
    &\qquad =
    \theta \int_{\IR^d} (r(y, \eta) g(y, u) - r(x, \eta) g(x, u)) q(x, dy)
    - \theta \int_{\IR^d} g(x, u) (r(y, \eta) - r(x, \eta)) q(x, dy) \\
    &\qquad \to
    \DG(gr)(x) - g(x) \DG r(x) \qquad \text{as }N \to \infty.
\end{align*}
i.e., to
$\DG\left(g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)
- g(\cdot, u) \DG\left(r(\cdot, \smooth{r}\eta(\cdot))\right)(x)$.
Furthermore, by our assumption on $g$,
$\int_u^N g(x, u_1) du_1 / N \to 1$ for all $x$ and $u$,
so the first term here converges to
\begin{align*}
    \gamma(x, \eta)
        \frac{
            \DG\left( g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)
            -
            g(\cdot, x) \DG\left( r(\cdot, \smooth{r}\eta(\cdot))\right)(x)
        }{
            g(x, u)
        } .
\end{align*}
This corresponds to each particle moving according to the motion induced by
$g \mapsto \gamma \DG(gr) - g \DG r$.
\comment{TODO: explain the $r$-tilted $\DG$ motion somewhere earlier, introducing notation for it.}
As for the second term,
since $\int g(y, u_1) q(x, dy) \to g(x, u_1)$
and $\theta/N \to \alpha$,
it converges to
\begin{align*}
    2 \alpha
    \gamma(x, \eta)
    r(x, \eta) 
    \int_u^\infty
    \left( g(x, u_1) - 1 \right)
    du_1
\end{align*}
This corresponds to births at higher levels at rate $2 \alpha \gamma r$.

The remaining term is \eqref{eqn:level_generator}:
\begin{align*}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \theta
        N^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align*}
The first part only uses the fact that $\theta/N \to \alpha$,
while as noted in \eqref{eqn:b_limit}, $b$ converges to $\gamma \DG r + F$,
so in total this term converges to
\begin{align*}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \alpha
        \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        \left\{
            \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
        \right\} u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align*}
In other words, the level of a particle at $x$ evolves according to
\begin{align*}
    \dot u
    =
    \alpha
        \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        \left\{
            \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
        \right\} u .
\end{align*}

\subsection{Generators for the spatial processes}
Before defining the generator of the lookdown representations, 
we recall the generator of the spatial processes.

We first define our pre-limit
$(\eta^{N}(t))_{t>0}$ as a solution
to a Martingale Problem
with generator $\Pgen^{N}$.

\begin{definition}
    \label{def2: MP definition of pre-limit}
Let $F \in \mathcal{C}^{2}(\IR)$ and $f \in \mathcal{C}_{0}(\IR^d)$.
We define $\Pgen^N$ to be a generator
acting on test functions of the form
$F_f(\eta):=F (\langle f, \eta \rangle)$
such that 
\begin{equation}
\begin{aligned}
\Pgen^{N} F_f(\eta):=& \int_{\IR^d} \theta\gamma(x, \eta) \left\{\int_{\IR^d} \left\{F(\langle f, \eta \rangle + f(y)/N )-F(\langle f, \eta \rangle )\right\}r(y, \eta)q_{\theta}(x,dy)\right\}N\eta(dx)\\
&+\int_{\IR^d} \theta \mu_\theta(x, \eta) \left\{F(\langle f, \eta \rangle - f(x)/N )-F(\langle f, \eta \rangle )\right\}N\eta(dx).
\end{aligned}    
\end{equation}
In this case, our pre-limit model $(\eta^{N}(t))_{t \geq 0}$ is the solution to the Martingale Problem $(\Pgen^{N}, \eta_0)$, i.e. 
for all $F \in \mathcal{C}^{2}(\IR)$
and $f \in \mathcal{C}_{0}(\IR^d)$,
$$M_t:=F_f(\eta^{N}_t)-F_f(\eta_0)
-\int_{0}^{t}\Pgen^{N}F_f(\eta^{N}_s)ds$$
is a martingale, and the initial condition is distributed as a Poisson measure with mean measure $\eta_0$.
\end{definition}

\begin{remark}
Taking $F(x)=x$ and $F(x)=x^2$, we can recover Equation \eqref{eqn:limiting_mgale_problem} and \eqref{eqn:limiting_mgale_variation} for the limit.
Similarly, we can recover
Definition \ref{defn:mgale_construction},
an equivalent characterisation
of our pre-limit model.
\end{remark}

We now write the candidate generator for the limiting object.

\begin{definition}
    \label{def2: MP definition of limit}
Let $F \in \mathcal{C}^{2}(\IR)$ and $f \in \mathcal{C}_{0}(\IR^d)$.
We define $\Pgen$ to be a generator
acting on test functions of the form
$F_f(\eta):=F (\langle f, \eta \rangle)$
such that 
\begin{equation}
    \label{eq: Limit Generator Definition}
\begin{aligned}
\Pgen^{\infty} F_f(\eta):=& F'(\langle f, \eta \rangle)
                   \times \left\{
                   \big\langle
                        \gamma(x, \smooth{\gamma} \eta(x))
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \smooth{r} \eta(\cdot))
                            \right)(x)
                    +
                    f(x)
                        F(x, \smooth{F} \eta(x)),
                        \eta(dx)
                    \big\rangle
                   \right\}\\
                &+ \alpha F''(\langle f, \eta \rangle)
                  \times \left\{
                  \big\langle
                    \gamma \left( x, \eta \right)
                    r\left(x,\eta \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle 
                  \right\}.
\end{aligned}    
\end{equation}
We define $(\eta^{\infty}(t))_{t \geq 0}$ to be the solution to the Martingale Problem $(\Pgen^{\infty}, \eta_0)$, i.e. 
for all $F \in \mathcal{C}^{2}(\IR)$
and $f \in \mathcal{C}_{0}(\IR^d)$,
$$M_t:=F_f(\eta^{\infty}_t)-F_f(\eta_0)
-\int_{0}^{t}\Pgen^{\infty}F_f(\eta^{\infty}_s)ds$$
is a martingale, with initial condition distributed as a Poisson measure with mean measure $\eta_0$.
\end{definition}

\subsubsection{Convergence of the Generators for the Spatial Models}

\comment{This is a duplicate with Lemma~\ref{lem:limit_mgale}; compare and remove this one.}

\begin{lemma}[Identifying the limit as a solution to the martingale problem]
    \label{lem:limit_mgale2}
As $N \to \infty$,
$(\eta^{N}_t)_{t \geq 0}$ converges in $\mathcal{D}_{[0,\infty)}(\mathcal{M}_f(\IR^d))$
to $(\eta^{N}(t))_{t \geq 0}$.
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:limit_mgale2}
We use Theorem 4.8.10 in \cite{EK}.
First of all, the set of functions
$\{F_f(\eta):= F(\langle f, \eta \rangle ),~
F \in \mathcal{C}^{2}(\IR), ~
f \in \mathcal{C}_{0}(\IR^d)\}$
is a dense and convergence-determining 
subset of functions in $\mathcal{M}_f(\IR^d)$.
Therefore, it suffices to show that for any $t>0$,
\begin{equation}
    \label{eq: Convergence Condition}
\lim_{N \to \infty}
\mathbb{E}\left[
\left(
F_f(\eta^{N}_{t+\tau})-F_f(\eta^{N}_t)
-\int_{t}^{t+\tau}\Pgen^{\infty}F_f(\eta^{N}_s)ds
\right)
\prod_{i=1}^{k}\langle h_i,\eta^{N}_{t_i} \rangle
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+\tau$,
and $h_1,...,h_k \in \mathcal{C}_{0}(\IR^d)$.

To prove this,
we make use of the Martingale Problem characterisation
of $(\eta^{N}_t)_{t \geq 0}$.
From Definition \ref{def: MP definition of pre-limit},
we can apply Tower Law
and obtain that
\begin{equation}
    \label{eq: Prelimit MP Application}
\lim_{N \to \infty}
\mathbb{E}\left[
\left(
F_f(\eta^{N}_{t+\tau})-F_f(\eta^{N}_t)
-\int_{t}^{t+\tau}\Pgen^{N}F_f(\eta^{N}_s)ds
\right)
\prod_{i=1}^{k}\langle h_i,\eta^{N}_{t_i} \rangle
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+\tau$,
and $h_1,...,h_k \in \mathcal{C}_{0}(\IR^d)$.

Therefore, it suffices to show that 
\begin{equation}
    \label{eq: Convergence of Spatial Generator}
\lim_{N \to \infty}
\mathbb{E}\left[
\int_{t}^{t+\tau}
\left|
\Pgen^{N}F_f(\eta^{N}_s)
-\Pgen^{\infty}F_f(\eta^{N}_s)
\right|
ds
\times 
\prod_{i=1}^{k}\langle h_i,\eta^{N}_{t_i} \rangle
\right]=0
\end{equation}
We apply Taylor Expansion
around the term $F(\langle f, \eta \rangle )$
up to third order.
This gives
\begin{equation}
    \label{eq: Taylor Expansion calculations i}
\begin{aligned}
&F(\langle f, \eta \rangle + f(y)/N )-F(\langle f, \eta \rangle )\\
=& F'(\langle f, \eta \rangle ) \frac{f(y)}{N}
+\frac{1}{2}F''(\langle f, \eta \rangle ) \frac{f^2(y)}{N^2}
+\frac{1}{6}F'''(w ) \frac{f^3(y)}{N^3}
\end{aligned}
\end{equation}
for some $w \in [\langle f, \eta \rangle, \langle f, \eta \rangle+ ||f||/N]$.
Similarly,
\begin{equation}
    \label{eq: Taylor Expansion calculations ii}
\begin{aligned}
&F(\langle f, \eta \rangle - f(x)/N )-F(\langle f, \eta \rangle )\\
=& - F'(\langle f, \eta \rangle ) \frac{f(x)}{N}
+\frac{1}{2}F''(\langle f, \eta \rangle ) \frac{f^2(x)}{N^2}
-\frac{1}{6}F'''(v ) \frac{f^3(x)}{N^3}
\end{aligned}
\end{equation}
for some $v \in [\langle f, \eta \rangle - ||f||/N, \langle f, \eta \rangle ]$.

We may now rewrite $\mathcal{P}^NF(\langle f, \eta \rangle)$ as 
\small
\begin{equation} 
    \label{eq: Pre-Limit Generator Expanded}
\begin{aligned}
\Pgen^{N} F_f(\eta):=& F'(\langle f, \eta \rangle)\int_{\IR^d}  \theta\left\{\int_{\IR^d} \left\{f(y)\gamma(x, \eta) r(y,\eta) -f(x)\mu_\theta(x, \eta)\right\}q_{\theta}(x,dy)\right\}\eta(dx)\\
&+\frac{1}{2}\frac{\theta}{N}F''(\langle f, \eta \rangle)\int_{\IR^d} 
 \left\{\gamma(x, \eta)\int_{\IR^d} \left\{f^2(y) r(y,\eta) \right\}q_{\theta}(x,dy)+f^2(x)\mu_\theta(x, \eta)\right\}\eta(dx)\\
&+\frac{1}{6}\frac{\theta}{N^2}\int_{\IR^d}\left\{ 
F'''(w) \left\{\gamma(x, \eta)\int_{\IR^d} f^3(y) r(y,\eta) q_{\theta}(x,dy)\right\}-F'''(v)f^3(x)\mu_\theta(x, \eta)\right\}\eta(dx)
\end{aligned}    
\end{equation}
\normalsize
for some $w,v \in [\langle f,\eta \rangle - ||f||/N, \langle f,\eta \rangle + ||f||/N]$.

As $N\to \infty$, 
$\theta/N \to \alpha$, 
$\theta/N^2 \to 0$,
and 
$$\int_{\IR^d} f^m(y) r(y,\eta) q_{\theta}(x,dy) \to f^m(x)r(x, \eta)$$
for $m=2,3$,
the third term in Equation \eqref{eq: Pre-Limit Generator Expanded} vanishes
while the second term converges to
that of Equation \eqref{eq: Limit Generator Definition}.
Furthermore,
$\mu_\theta(x,\eta) = r(x,\eta)\gamma(x,\eta)-\frac{1}{\theta}F(x,\eta) \to r(x,\eta)\gamma(x,\eta),$
so $f^2(x)\mu_\theta(x,\eta) \to r(x,\eta)\gamma(x,\eta) f^2(x)$.

Finally, by Equation \eqref{mean measure},
\eqref{eqn:rewritten mean measure}, \eqref{eqn:near_critical} and
\eqref{limit of mean measure equation},
we have 
\begin{multline}
 \theta\left\{\int_{\IR^d} \left\{f(y)\gamma(x, \eta) r(y,\eta) -f(x)\mu_\theta(x, \eta)\right\}q_{\theta}(x,dy)\right\}
 \\
\to \gamma(x, \eta)                           \mathcal{B}\left(
                            f(\cdot) r(\cdot, \eta)
                            \right)(x)
                    +
                    f(x)
                        F(x, \eta),
\end{multline}
so comparing terms we have
\begin{equation}
\lim_{N\to \infty} |\mathcal{P}^{N}F\langle f, \eta \rangle - \mathcal{P}^{\infty}F\langle f, \eta \rangle| \to 0,   
\end{equation}
and so by Dominated Convergence Theorem,
we establish Equation 
\eqref{eq: Convergence of Spatial Generator}
which concludes our proof.
\end{proof}

\subsubsection{Exact Statement on Markov Mapping Theorem results}

\begin{theorem}
\label{teo: lookdown on pre-limit well defined}
For each $N \in \mathbb{N}$,
there exists a solution $(\lp^N_t)_{t \geq 0}$ to the $(A^N, \lp^N_0)$ martingale problem.
Furthermore, if we define $\Xi^N_t = \frac{1}{N}\lp^N_t(dx \times [0,N])$ for $t \geq 0$,
then the pre-limiting model $(\eta^N_t(dx))_{t \geq 0}$ defined in Definition \ref{defn:mgale_construction}
and $(\Xi^N_t(dx))_{t \geq 0}$
shares the same distribution on $\mathcal{D}_{[0,\infty)}(\mathcal{M}(\mathbb{R}^d))$.

Finally, 
for all $h\in \mathcal{C}(\mathbb{R}^d \times [0, N])$
\begin{equation}\label{Markov Mapping Theorem Cox Condition}
\mathbb{E}\left[e^{-\int_{\mathbb{R}^d \times [0,N]} h(x,u) \xi^{N}(dx \times du)}\bigg| \mathcal{F}^{\Xi^{N}}_t\right]=e^{-\int F^{N}_h(x) \Xi^{N}_t(dx)},
\end{equation}
where 
$$F^{\epsilon,\theta,N}_h(x):=-N\log \left(\frac{1}{N}\int_{0}^{N}e^{-h(x,u)}du\right)=-N\log \left(1-\frac{1}{N}\int_{0}^{N}(1-e^{-h(x,u)})du\right).$$
In other words,
conditioned on the spatial configuration $\Xi^N_t$,
the levels of the particles must be i.i.d uniformly between $[0,N]$.

\end{theorem}

The statement for the limiting case is similar,
with the only difference being the conditional distribution 
of the levels given the spatial projection.

\begin{theorem}
\label{teo: lookdown on limit well defined}
There exists a solution $(\lp_t)_{t \geq 0}$ to the $(A, \lp_0)$ martingale problem.

Furthermore, if we define $\Xi_t = \lim_{u_0 \to \infty} \frac{1}{u_0}\lp_t(dx \times [0,u_0])$,
then the limit model $(\eta_t(dx))_{t \geq 0}$ defined in Theorem \ref{thm:nonlocal_convergence}
and $(\Xi_t(dx))_{t \geq 0}$
shares the same distribution on $\mathcal{D}_{[0,\infty)}(\mathcal{M}(\mathbb{R}^d))$.

Finally, 
for all $h\in \mathcal{C}(\mathbb{R}^d)$ and for all $K>0$,
\begin{equation}\label{Markov Mapping Theorem Cox Condition limit}
\mathbb{E}\left[e^{-\int_{\mathbb{R}^d \times [0,K]} h(x) \xi(dx \times du)}\bigg| \mathcal{F}^{\Xi}_t\right]=e^{- K \int_{\mathbb{R^d}}(1-e^{h(x)}) \Xi_t(dx)}.
\end{equation}
In other words,
conditionally on $(\Xi_s)_{0\leq s \leq t}$
for each $s \in[0,t]$,
$\xi_s$ is conditionally Poisson with Cox measure $\Xi_s(dx) \times du$.
\end{theorem}

\subsection{Markov Mapping Theorem Application on Generator $A^{N}$}\label{sec: Markov Mapping Theorem Application - Prelimit}

We now introduce some notations to simplify our calculations.
For a finite atomic measure $\xi=\sum \delta_{(x_i,u_i)} \in \mathcal{M}(\mathbb{R}^d \times [0,N])$,
we introduce the test functions
 $$f_g(\xi)= \prod_{(x_i,u_i)\in \xi} g(x_i,u_i) = \exp (\langle \log g, \xi \rangle),$$
 where $g \in \mathcal{C}(\mathbb{R}^d \times [0,\infty))$ and $g(x,u)=1$ for all $u \geq N$.

>>>>

The lookdown processes have been carefully constructed so that
observations about the past spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels.
In other words, the dynamics of the lookdown
process preserve the conditionally uniform (or in the limit, conditionally Poisson) structure
-- if started with uniform levels,
levels are uniform at all future times.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
the Markov Mapping Theorem (Theorem~\ref{thm:mmt}; also see \citep{etheridge/kurtz:2019})
tells us that by ``removing labels'' from the lookdown process $\lp$
we recover the population process $\eta$.
% In particular, for any
% solution to the martingale problem of Definition~\ref{defn:mgale_construction}
% % of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
% there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
% solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
% for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
% and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
% are independently and uniformly distributed on $[0, N]$.
% Similarly, in the scaling limit:
% for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
% Theorem~\ref{nonlocal_convergence}
% %there is a lookdown process $(\lp^N_t)_{t \ge 0}$
% there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
% for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
% and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
% $\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
% Furthermore, convergence of the lookdown processes would imply convergence of the population processes
% \citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

@@

\begin{proof}[of Proposition~\ref{thm:mmt_application}]
    This proposition is the content of the Markov Mapping Theorem,
    reproduced from \citet{etheridge/kurtz:2019} as Theorem~\ref{thm:mmt},
    applied to our situation.
    In the notation of that theorem, $\gamma$ is what we have called $\kappa$ above,
    $\alpha$ is the transition function that assigns levels uniformly on $[0, N]$ (in the first case)
    or as a Poisson process with Lesbegue intensity (in the limiting case),
    and we need a continuous $\psi^N(\lp)$ such that $|A^N f(\lp)| \le c_f \psi^N(\lp)$
    for all $f$ in the domain of $A^N$ (and similarly a function $\psi$ for $A$).
    We also need that applying the lookdown generator to a function and averaging over levels
    is equivalent to applying the population process generator to the function
    whose dependence on levels has been averaged out,
    a condition which we precisely state, and verify,
    in Lemmas~\ref{lem:averaged_generators} and~\ref{lem:averaged_generators_limit}
    of the Appendix.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C (1 + u |F(x, \eta)|), \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the most troublesome term in $Af(\lp)$,
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align*} % \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align*}
    The first line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the second line provides us with the required bound:
    we absorb $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and take $\psi(\lp) = \langle F(x, \eta) u e^{-u}, \lp(dx, du) \rangle$.
    % Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    % for a given $\lp$ (and hence $\eta = \kappa(\lp)$),
    % $\sup_x |F(x, \eta)|$ is finite.
    % Therefore, we may bound \eqref{eqn:part_of_psi} by
    % $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    % To do this we must restrict $\lp$ to lie in the space of measures
    % for which this is finite;
    % however, this is a completely metrizable space
    % \citep[see Condition 2.1 in][]{etheridge/kurtz:2019},
    % and our processes fall in this space (by conditional uniformity).
\end{proof}

@@

% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %%
\section{Markov Mapping Theorem}
\label{apx:mmt}

The following appears as Theorem A.2 in \citet{etheridge/kurtz:2019},
specialized slightly to the case that the processes 
are c\'adl\'ag and have no fixed points of discontinuity.
For an $S_0$-valued, measurable process $Y$, $\hat{\mathcal{F}}^Y_t$
denotes the completion of the $\sigma$-algebra generated by
$Y(0)$ and $\{\int_0^r h(Y(s)) ds, r \le t, h \in B(S_0)\}$.
Also, let $D_S[0,\infty)$ denote the space of c\`adl\`ag, $S$-valued functions
with the Skorohod topology, and $M_S[0,\infty)$ the space of Borel measurable functions
from $[0,\infty)$ to $S$,
topologized by convergence in Lesbegue measure.
For other definitions see \citet{etheridge/kurtz:2019}.

\begin{theorem}[Markov Mapping Theorem] \label{thm:mmt}
    Let $(S,d)$ and $(S_0,d_0)$ be complete, separable metric spaces.
    Let $A \subset \bar C(S) \times C(S)$ and $\psi \in C(S)$, $\psi \ge 1$.
    Suppose that for each $f \in \mathcal{D}(A)$ there exists $c_f$
    such that
    $$ |Af(x)| \le c_f \psi(x), \qquad x \in A, $$
    and define $A_0 f(x) = Af(x) / \psi(x)$.

    Suppose that $A_0$ is a countably determined pre-generator,
    and suppose that $\mathcal{D}(A) = \mathcal{D}(A_0)$
    is closed under multiplication and is separating.
    Let $\gamma : S \to S_0$ be Borel measurable,
    and let $\alpha$ be a transition function from $S_0$ into $S$
    ($y \in S_0 \to \alpha(y,\cdot)\in\mathcal{P}(S)$ is Borel measurable)
    satisfying $\int h\circ\gamma(x)\alpha(y,dx) = h(y)$ for $y \in S_0$ and $h \in B(S_0)$,
    that is, $\alpha(y,\gamma^{-1}(y)) = 1$.
    Assume that $\tilde \psi(y) \equiv \int_S \psi(z) \alpha(y,dz) < \infty$ for each $y \in S_0$
    and define
    $$  C = \{
        \int_S f(z) \alpha(\cdot,dz),
        \int_S Af(z) \alpha(\cdot,dz)
        \; : \; f \in \mathcal{D}(A)
    \} . $$
    Let $\mu_0 \in \mathcal{P}(S_0)$ and define $\nu_0 = \int \alpha(y,\cdot) \mu_0(dy)$.
    \begin{itemize}
        \item[(a)]
            If $\tilde Y$ satisfies $\int_0^t \IE[\tilde \psi(\tilde Y(s))]ds < \infty$
            for all $t \ge 0$ and $\tilde Y$ is a solution of the martingale problem for $(C,\mu_0)$,
            then there exists a solution $X$ of the martingale problem for $(A,\nu_0)$
            such that $\tilde Y$ has the same distribution on $M_{S_0}[0,\infty)$
            as $Y = \gamma \circ X$.
            If $Y$ and $\tilde Y$ are c\'adl\'ag, then $Y$ and $\tilde Y$ have the same distribution
            on $D_{S_0}[0,\infty)$.
        \item[(b)]
            For $t \ge 0$,
            $$ \IP\{ X(t) \in \Gamma \;|\; \hat{\mathcal{F}}^Y_t \}
            = \alpha(Y(t),\Gamma), \qquad \text{for } \Gamma \in \mathcal{B}(S).
            $$
        \item[(c)]
            If, in addition, uniqueness holds for the martingale problem for $(A,\nu_0)$,
            then uniqueness holds for the $M_{S_0}[0,\infty)$-martingale problem for $(C,\mu_0)$.
            If $\tilde Y$ has sample paths in $D_{S_0}[0,\infty)$ then uniqueness holds for the
            $D_{S_0}[0,\infty)$-martingale problem for $(C,\mu_0)$.
        \item[(d)]
            If uniqueness holds for the martingale problem for $(A,\nu_0)$
            then $Y$ is a Markov process.
    \end{itemize}
\end{theorem}

In our application, we have taken $S$ to be the space of locally finite counting measures
on $\IR^d \times [0,N)$ or on $\IR^d \times [0,\infty)$,
and $S_0$ the space of finite measures on $\bar \IR^d$.
Then, $A$ corresponds to the generator for the lookdown process (i.e., either $A^N$ or $A$),
and $C$ corresponds to the generator for the spatial population process
(i.e., either $\Pgen^N$ or $\Pgen$).
The ``$\gamma$'' of the theorem is our spatial projection operator
that we have called $\kappa^N$ or $\kappa$,
and the ``$\alpha$'' of the theorem will be named $\Gamma^N$ or $\Gamma$ below.
Finally, ``$X$'' of the theorem is our lookdown process, $\lp$,
and ``$Y$'' is our spatial process, $\eta$.

@@
